---
title: "MSDS6306 CaseStudy 2"
date: "December 4, 2017"
output:
    html_document:
        keep_md: true
---

## 1. Project Goals 

###Identify (at least) the top three factors that contribute to turnover.
###Learn about any job role specific trends that may exist in the data set

```{r,echo=FALSE, message=FALSE}
require(xlsx)
require(leaps)
require(MASS)
require(tidyverse)
require(reshape2)
require(ggplot2)
require(ggmosaic)
require(plyr)
#require(psych)
```

## 2. import data 
```{r,echo=FALSE, message=FALSE}
employee<-read.xlsx('C://Git//Repository//DS6306_casestudy2//data//CaseStudy2-data.xlsx',sheetIndex = 1)
```

## 3. Initial Data Exploration 

###  3.1 Check for NAs and Nulls  
### - there is no NAs or Nulls in the dataset.
```{r,echo=TRUE}
unlist(lapply(employee, function(x) any(is.na(x))))
unlist(lapply(employee, function(x) any(is.null(x))))
```


###  3.2 Initial inspect of all columns. Drop irrelevant columns that has only 1 level of values. 
```{r,echo=TRUE}
employee<-employee[,!(names(employee) %in% c("Over18","EmployeeCount","StandardHours"))]
```

### 3.3  Check factor column,Add numeric columns for factor columns for building correlation matric and regression model 

```{r,echo=TRUE}
# ename<-names(employee)
fnames<-names(Filter(is.factor, employee))

employee$BusinessTravelN<-as.numeric(employee$BusinessTravel)
employee$OverTimeN<-as.numeric(employee$OverTime)
employee$DepartmentN<-as.numeric(employee$Department)
employee$EducationFieldN<-as.numeric(employee$EducationField)
employee$GenderN<-as.numeric(employee$Gender)
employee$OverTimeN<-as.numeric(employee$OverTime)
employee$MaritalStatusN<-as.numeric(employee$MaritalStatus)
employee$JobInvolvementN<-as.numeric(employee$JobInvolvement)
employee$JobRoleN<-as.numeric(employee$JobRole)
employee$AttritionN<-as.numeric(employee$Attrition)
employee$AgeGroups <- cut(employee$Age, breaks = 5)
```
### 3.4 Build Correlation matrix for all numeric variables 
### Get an initially look of the data and all correlated variables
```{r,eval=FALSE,Echo=TRUE}
## Fix an issue that existed in psych package that might impact ggplot function.
 #devtools::install_github("Artjom-Metro/ggsignif")
 nnames<-names(Filter(is.numeric,employee))
 #lowerCor(employee[,nnames])
```

## 4. Attrition Analysis. 

### 4.1 Run automatic methods for variable selection from the choosen variables. 

```{r, Echo=TRUE}
regBest<- regsubsets(Attrition ~ Age + BusinessTravel + DailyRate +Department  + DistanceFromHome + 
                     EducationField + JobLevel+ JobRole +JobSatisfaction +MaritalStatus +
                     MonthlyIncome + NumCompaniesWorked + OverTime + TotalWorkingYears + 
                     TrainingTimesLastYear + YearsAtCompany + YearsInCurrentRole +
                     YearsSinceLastPromotion + YearsWithCurrManager
                    ## Variables Need further inspection 
                    + EmployeeNumber + EnvironmentSatisfaction + StockOptionLevel
                    + StockOptionLevel + YearsSinceLastPromotion 
                    + Education + Gender + HourlyRate + JobInvolvement
                    + MonthlyRate + PerformanceRating + RelationshipSatisfaction
                    + WorkLifeBalance
                    ## Variables unnessesary
                    ## + Over18 + EmployeeCount + StandardHours
                    ,
                  data = employee, nvmax = 5) 

# summary(regBest)
plot(regBest,col = "brown",main = "Variables chosen by best selection model")

regBack<- regsubsets(Attrition ~ Age + BusinessTravel + DailyRate +Department  + DistanceFromHome + 
                       EducationField + JobLevel+ JobRole +JobSatisfaction +MaritalStatus +
                       MonthlyIncome + NumCompaniesWorked + OverTime + TotalWorkingYears + 
                       TrainingTimesLastYear + YearsAtCompany + YearsInCurrentRole +
                       YearsSinceLastPromotion + YearsWithCurrManager
                     ## Variables Need further inspection 
                     + EmployeeNumber + EnvironmentSatisfaction + StockOptionLevel
                     + StockOptionLevel + YearsSinceLastPromotion
                     + Education + Gender + HourlyRate + JobInvolvement
                     + MonthlyRate + PerformanceRating + RelationshipSatisfaction
                     + WorkLifeBalance
                     ## Variables unnessesary
                     ## + Over18 + EmployeeCount + StandardHours 
                     ,
                     data = employee, nvmax = 5, method = "backward") 
#summary(regBack)
plot(regBack,col = "brown",main = "Variables chosen by backward selection model")
```
### From the backward selection model,top 5 Variables impacting Attrition the most are: 
### OverTime(Yes),MaritalStatus(Single), Age,EnvironmentSatisfaction,JobRole(Sales)
### From the best selection model,top 5 Variables impacting Attrition the most are: 
### OverTime(Yes),MaritalStatus(Single),TotalWorkingYears,EnvironmentSatisfaction,JobInvolvement

### Therefore, we would choose OverTime, MaritalStatus, EnvironmentSatisfaction, TotalWorkingYears, Age, JobInvolvement and JobRole for further analysis.

### 4.2 Further variance inspection 

```{r, Echo=TRUE}
##<<<<<<< HEAD

#### Probabilities of value given attrition #### 

attrified <- employee[employee$Attrition == 'Yes',]
not.attrified <- employee[employee$Attrition == 'No', ]

Attrition_prop_table <- function(variable_name, data.f){
    # Generates a table containing proportion of responses for both Attrition values. This should allow us to examine values in the context of whether they attrified.
    
    # Generate a table containing variable/Attrition rates
    prop <-prop.table(xtabs(as.formula(paste( '~ ',paste(variable_name, 'Attrition ', sep = ' + '))) , data=data.f))
    
    #Normalize each column to sum to 1.
    prop.app <-apply(prop,2,sum)
    return(melt(sweep(prop, MARGIN=2,prop.app,'/')))
}


EnvSatTest_attrit <- Attrition_prop_table('EnvironmentSatisfaction',employee)
MariStat_attrit <- Attrition_prop_table( 'MaritalStatus',employee)
```
### Employees that left are most likely to be single.
```{r, Echo=TRUE}
OverTime_attrit <- Attrition_prop_table( 'OverTime',employee)
JobInvolv_attrit <- Attrition_prop_table( 'JobInvolvement',employee)
JobRole_attrit <- Attrition_prop_table( 'JobRole',employee)
```
### We have highest relative turnover of Sales Representative. Very few Research Directors leave.
```{r, Echo=TRUE}
AgeGroup_attrit <- Attrition_prop_table( 'AgeGroups', employee)
```
### Younger employees are more likely to leave.
```{r, Echo=TRUE}
table(employee[employee$Age < 26.4,'JobRole']) # Lab Tech, Research Scientist, 
table(employee[employee$Age < 26.4 & employee$Attrition =='Yes','JobRole']) 
```
### Sales Rep, Lab Tech, Research Scientist are the jobs most often left.
```{r, Echo=TRUE}
##<<<<<<< HEAD
## 100% Stacked Bar chart
ggplot(EnvSatTest_attrit, aes(x=Attrition, y=value, fill= factor(EnvironmentSatisfaction))) + 
    geom_bar(position = "fill",stat = "identity") +
    scale_y_continuous(breaks = c(1,2,3,4)) +
    labs(title = 'Attrition rate by\nEnvironment Satisfaction', y='Proportion of Response') +
    guides(fill=guide_legend(title="Environment \nSatisfaction"))+
    theme(legend.title=element_text(size = 9),plot.title=element_text(hjust=0.5))
```

### 4.3 We'll go ahead and create multiple linear regression model to get more evidence in order to determine which factors imapct Atrrition the most. 

### Build regression model
```{r, Echo=TRUE}
mylm<-lm(AttritionN ~ OverTimeN + MaritalStatusN + EnvironmentSatisfaction + TotalWorkingYears+ Age + JobInvolvementN + JobRoleN, data = employee)
summary(mylm)
```
### factors significant are:  OverTimeN,MaritalStatusN,EnvironmentSatisfaction,JobInvolvementN,TotalWorkingYears and Age.
### Select top 3 factors that contribut to attrition the most
### The top 3 factors are: OverTime,EnvironmentSatisfaction,MaritalStatus, 

### 4.4 Combine values in the factors for further analysis. 
```{r, Echo=TRUE}
employee$MaritalStatusR<-as.character(employee$MaritalStatus)
employee$MaritalStatusR[employee$MaritalStatusR!="Single"]<-c("Not Single")
employee$EnvironmentSatisfactionR<-employee$EnvironmentSatisfaction
employee$EnvironmentSatisfactionR[employee$EnvironmentSatisfaction!=1]<-c("Medium or Higher Satisfation")
employee$EnvironmentSatisfactionR[employee$EnvironmentSatisfaction==1]<-c("Low Satisfation")
```
### 4.5 Plot the data 
```{r,Echo=TRUE}
bg<-ggplot(employee, aes(x=OverTime,fill=Attrition))
bg<-bg + geom_bar(position = "fill")
bg<-bg + scale_y_continuous(labels = scales::percent)
bg<-bg + labs(y="Percentage", x="Over Time", title="Attrition vs OverTime")
bg

bg<-ggplot(employee, aes(x=MaritalStatusR,fill=Attrition))
bg<-bg + geom_bar(position = "fill")
bg<-bg + scale_y_continuous(labels = scales::percent)
bg<-bg + labs(y="Percentage", x="MaritalStatus", title="Attrition vs MaritalStatus")
bg

bg<-ggplot(employee, aes(x=EnvironmentSatisfactionR,fill=Attrition))
bg<-bg + geom_bar(position = "fill")
bg<-bg + scale_y_continuous(labels = scales::percent)
bg<-bg + labs(y="Percentage", x="Environment Satisfaction", title="Attrition vs Environment Satisfaction")
bg
```
### 4.6 Plot masaic plot for combing multiple factor's effection 

#### From the analysis, we can get the top 3 factors that contribuilt to a higher turnover rate of an employee:
#### 1. Working overtime.
#### 2. Being single.
#### 3. Low satisfaction of working enviorment.

```{r, Echo=TRUE}
bg<-ggplot(employee)
bg<-bg + geom_mosaic(aes( x = product(MaritalStatusR,OverTime,EnvironmentSatisfactionR),
                          fill=Attrition)) 
bg<-bg + theme(axis.text.x=element_text(angle=35, hjust= 1))
bg<-bg + labs(x="Marital Status, Eviroment Satisfaction", title='Emploree attrition vs Overtime/MaritalStatus/EnvironmentSatisfaction',divider= "hspine")
bg<-bg + guides(fill=guide_legend(reverse = TRUE))
bg

mosaicplot(~Attrition + OverTime + MaritalStatusR + EnvironmentSatisfactionR, 
           shade = T, data = employee ,
           main = "Residules of Attrition vs Overtime/MaritalStatus/EnvironmentSatisfaction",
           las = 5,
           border = "chocolate",
           off = 10
)
```

## 5. Job Role Analysis. 

### 5.1 There are nine job roles in the dataset 
```{r, Echo=TRUE}
levels(employee$JobRole)
```

###  5.2 Initial factors Screen. Chose a meaningful factor for analysis: Numer of the Companies Worked. 

### 5.3 Create subset for Analysis 
```{r, Echo=TRUE}
sub.JRNCW<-employee[,c("NumCompaniesWorked","JobRole","JobRoleN")]
mean.JRNCW<-ddply(employee,~JobRole,summarise,meanNCW=mean(NumCompaniesWorked))
```

###  5.4 Build model and interpret the results 
```{r, Echo=TRUE}
lm1<-lm(formula = NumCompaniesWorked ~ JobRole, data = sub.JRNCW)
summary(lm1)
```
### Research Director and Sales representative shows significant difference in number of companies worked than other roles.
```{r, Echo=TRUE}
sub.JRNCW$JobRoleR<-as.character(sub.JRNCW$JobRole)
sub.JRNCW$JobRoleR[sub.JRNCW$JobRoleR!="Research Director"]<-c("Non Research Director")

sub.JRNCW$JobRoleS<-as.character(sub.JRNCW$JobRole)
sub.JRNCW$JobRoleS[sub.JRNCW$JobRoleS!="Sales Representative"]<-c("Non Sales")
```

###Significant evidence shows that the mean number of companies that a Research Director had ever worked is greater than other roles. 
###p-vlue=0.00017 at significant level alpha=0.05
### 95% confidence level of how many more companies a Reseach Director had every worked is 0.5 to 1.6.
```{r, Echo=TRUE}
aov1<-aov(NumCompaniesWorked~JobRoleR,data = sub.JRNCW)
summary(aov1)
confint(aov1)
```
###Significant evidence shows that the mean number of companies that a Sales Representative had ever worked is less than other roles. 
### p-vlue=<0.0001. at significant level alpha=0.05
### 95% confidence level shows: The companies that a Sales Representative have ever worked is 0.5 to 1.7 less than other groups of roles.
```{r, Echo=TRUE}
aov2<-aov(NumCompaniesWorked~JobRoleS,data = sub.JRNCW)
summary(aov2)
confint(aov2)

```
###  5.5 Conclusion:  
### From the analysis, Sales Representative workes less companies than other groups of roles and Research Director works more companies than other group of roles.
### But this results might be impacted by confound variances. Such as , Sales Representative's Age, working years and so on. Further analysis need to proceed on the test to determine confound variables.
```{r, Echo=TRUE}
bg<-ggplot(sub.JRNCW, aes())
bg<-bg + geom_bar(aes(x=JobRoleR, y=NumCompaniesWorked, fill = JobRoleR), 
                  position = "dodge", stat = "summary", fun.y = "mean")
bg
bg<-ggplot(sub.JRNCW, aes())
bg<-bg + geom_bar(aes(x=JobRoleS, y=NumCompaniesWorked, fill = JobRoleS), 
                  position = "dodge", stat = "summary", fun.y = "mean")
bg
```